---
title: "clean_merge"
output: html_document
date: "`r Sys.Date()`"
---

load libraries 

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(lubridate)

```

Load up data

```{r}

Data <- read.csv("/home/aabdulrasul/Documents/Projects/SPINR/SPN30IdentificationO_DATA_2024-09-05_1113.csv")

# Run any pre-processing steps here that make the data readable - e.g, turn blanks "" into NA

# Replace blank with NA

replace_blank_with_na <- function(x) {
  if (is.character(x)) {
    x <- na_if(x, "")
  }
  return(x)
}

Data <- Data %>% mutate(across(everything(), ~ replace_blank_with_na(.)))
  
```


### Merge


Data set has 4 arms, we will separate them into 4 separate data frames. The logic behind this is to clean and process each on before we merge them into one final data frame

```{r}

Arm1 <- Data %>% filter(grepl("_arm_1$", redcap_event_name)) # ASD
Arm2 <- Data %>% filter(grepl("_arm_2$", redcap_event_name)) # SSD in SPINS
Arm3 <- Data %>% filter(grepl("_arm_3$", redcap_event_name)) # Control
Arm4 <- Data %>% filter(grepl("_arm_4$", redcap_event_name)) # Repeat scan using some of the original study 
```

Filter out time_points in Arm4

```{r}

# Specifically in Arm4 we have instances where record_ids have separate rows for their respective time points. We want to ensure the merging process does not treat these instances as one. As such we are going to filter out these instances and save them in a new data frame _tx2 

Arm4_tx2 <- Arm4 %>% 
  filter(redcap_event_name == "tx2_arm_4")

Arm4 <- Arm4 %>%
  filter(redcap_event_name != "tx2_arm_4")

# For some reason Arm4 has a record_id named test - we will filter this out too

Arm4 <- Arm4 %>%
  filter(record_id != "test")
```



### Merge : Find Duplicates

Compare record_ids between the arms to determine if there are any duplicates.

```{r}

# First, extract the record_id columns from each dataframe
record_ids_arm_1 <- Arm1$record_id
record_ids_arm_2 <- Arm2$record_id
record_ids_arm_3 <- Arm3$record_id
record_ids_arm_4 <- Arm4$record_id

# Find common record_ids between each pair of arms
common_arm_1_2 <- intersect(record_ids_arm_1, record_ids_arm_2)
common_arm_1_3 <- intersect(record_ids_arm_1, record_ids_arm_3)
common_arm_1_4 <- intersect(record_ids_arm_1, record_ids_arm_4)
common_arm_2_3 <- intersect(record_ids_arm_2, record_ids_arm_3)
common_arm_2_4 <- intersect(record_ids_arm_2, record_ids_arm_4)
common_arm_3_4 <- intersect(record_ids_arm_3, record_ids_arm_4)

# Print the results
cat("Common IDs between Arm 1 and Arm 2:", common_arm_1_2, "\n")
cat("Common IDs between Arm 1 and Arm 3:", common_arm_1_3, "\n")
cat("Common IDs between Arm 1 and Arm 4:", common_arm_1_4, "\n")
cat("Common IDs between Arm 2 and Arm 3:", common_arm_2_3, "\n")
cat("Common IDs between Arm 2 and Arm 4:", common_arm_2_4, "\n")
cat("Common IDs between Arm 3 and Arm 4:", common_arm_3_4, "\n")
```


Filter out problematic duplicate record id : SPN30_CMH_043281 exists in both Arm2 and Arm4, however, Arm2 has a partially completed timepoint 1 dataset, this is causing some issues in the merge process as the same participant seemingly has completed the timepoint 1 battery in Arm4 which was completed a couple of days apart from Arm2. As such we can assume Arm4 is the more accurate dataset. therefore we will remove SPN30_CMH_043281 from Arm2.



```{r}
Arm2 <- Arm2 %>%
  filter(record_id != "SPN30_CMH_043281")
```


### Merge : Standardize duplicates

Now that we've determined Arm 2 and Arm 4 both have duplicates, we need to ensure that the information in these data frames are in fact duplicates of each other - or whether one has more data points filled than the other. We will essentially compare the 2 and then copy whatever is extra onto Arm 4 (as we are currently working with this data set).

Note that this is hard coded for Arm4 and Arm2, update and adapt where applicable based on the results of the previous cell


```{r}

# Identify common record_ids across Arm2 and 4
common_arm_2_4 <- intersect(Arm2$record_id, Arm4$record_id) 

# Filter both data frames to include only the common record IDs
filtered_Arm2 <- Arm2 %>% filter(record_id %in% common_arm_2_4)
filtered_Arm4 <- Arm4 %>% filter(record_id %in% common_arm_2_4)

# Merge the filtered data from Arm2 and 4 based on record_id
comparison_df <- merge(filtered_Arm2, filtered_Arm4, by = "record_id", suffixes = c("_arm2", "_arm4"))

# Compare each column and update directly
for (col in setdiff(names(filtered_Arm2), "record_id")) {
  col_arm2 <- paste0(col, "_arm2")
  col_arm4 <- paste0(col, "_arm4")
  
  # Compare and decide to update
  comparison_df[[col_arm4]] <- ifelse(comparison_df[[col_arm2]] != comparison_df[[col_arm4]] & !is.na(comparison_df[[col_arm2]]),
                                      comparison_df[[col_arm2]],
                                      comparison_df[[col_arm4]])
}

# Drop the arm2 columns and rename arm4 columns to their original names
updated_arm_4 <- comparison_df %>%
  select(record_id, ends_with("_arm4")) %>%
  rename_with(~ sub("_arm4", "", .), ends_with("_arm4"))


# Append these rows to updated_arm_4
complete_arm_4 <- bind_rows(updated_arm_4,  anti_join(Arm4, updated_arm_4, by = "record_id"))

```

(Optional) Remove intermediary data frames

```{r}
rm(comparison_df, filtered_Arm2, filtered_Arm4, updated_arm_4)
```


Now that we have the duplicates sorted we can now merge the arms together, note that we are going to merge Arm1, Arm2, Arm3 and Arm4. The steps above have effectively combined the duplicates seen in Arm2 and Arm4 under complete_arm_4 which means that when we merge the arms together, we need to ensure we are ONLY merging the non-duplicates from Arm2. All Arms will be merged as the dataframe MergedDF 

```{r}

# Merge all arms together into one dataframe. Anti_join is used here to find the unique record_ids between Arm2 and complete_arm_4 thereby allowing us to merge the non duplicates in Arm2, 

MergedDF <- bind_rows(Arm1, anti_join(Arm2, complete_arm_4, by = "record_id"), Arm3, complete_arm_4)

```

Some metrics for us to keep track of (in case we were wondering)

Note we have recruited 88 participants total (see recruitment.rmd)

```{r}

cat("Data collected from", sum(MergedDF$elg_form_complete == "1", na.rm = TRUE), "Eligible Participants\n")

```

## Neuropsych Assessments

The following Neuropsych assessments are collected, RMET (Reading The Mind in The Eyes Test), ER-40 (Penn Emotion Recognition Test), and, TASIT-R (The Awareness of Social Inference Test - Revised) Along with various clinical assessments. We will create a new dataframe combining all of these into one from MergedDF - this new dataframe will bec called Neuro

```{r}

 

```



